---
title: "P8105_hw2_PaulinaHan"
author: "Paulina Han"
date: "2021/10/1"
output: github_document
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
library(haven)
library(lubridate)
```

# Problem 1
```{r, warning=FALSE}
trash_df = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) 

#tidying trash data set
trash_tidy = 
  trash_df %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards) %>% 
  mutate(sports_balls = round(sports_balls,0))

#2019 precipitation
precipitation19_df = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet = "2019 Precipitation", range = "A2:B14")

#tidy 2019 precipitation data
tidy_2019_df =
  precipitation19_df %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2019)
  
#2018 precipitation
precipitation18_df = read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",sheet = "2018 Precipitation", range = "A2:B14")

#tidy 2018 precipitation data
tidy_2018_df =
  precipitation18_df %>% 
  janitor::clean_names() %>% 
  drop_na(total) %>% 
  mutate(year = 2018)  

#combine precipitation data set
precipitation = 
  bind_rows(tidy_2018_df, tidy_2019_df) %>% 
 mutate(month = month.name[month])

#description 
#data set 1
ncol_trash_tidy = ncol(trash_tidy)
nrow_trash_tidy = nrow(trash_tidy)

mean_weight = mean(pull(trash_tidy, weight))
max_weight = max(pull(trash_tidy, weight))
min_weight = min(pull(trash_tidy, weight))

mean_volume = mean(pull(trash_tidy, volume))
max_volume = max(pull(trash_tidy, volume))
min_volume = min(pull(trash_tidy, volume))

#data set 2
ncol_precipitation = ncol(precipitation)
nrow_precipitation = nrow(precipitation)

max_rain = max(pull(precipitation, total))

#median of sports ball in 2019
ball =
  trash_tidy %>% 
  filter(year == 2019) %>% 
  pull(sports_balls) %>% 
  median()

#total precipitation in 2018
sum_2018 = precipitation %>% 
  filter(year == '2018') %>% 
  pull(total) %>% 
  sum()

```

The tidied Mr.Wheel Trash data set has `r ncol_trash_tidy` columns and `r nrow_trash_tidy ` observations. The mean weight of trash the dumpster took is `r mean_weight` and the maximum weight of trash the dumpster took is `r max_weight`, the minimum weight of trash the dumpster took is `r min_weight`. The mean volume of trash the dumpster took is `r mean_volume`, the maximum volume of trash the dumpster took is `r max_volume` and the minimum volume of trash the dumpster took is `r min_volume`.

The precipitation data set including 2018 and 2019 has `r ncol_precipitation` columns and `r nrow_precipitation` observations. 2018 March had the most precipitation in 2018 and 2019 which is `r max_rain`.

The median number of sports balls in a dumpster in 2019 is `r ball`.

The total precipitation in 2018 is `r sum_2018`.

# Problem 2
```{r}
#read pols-month.csv
pols = read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/pols-month.csv")

#clean pols-month data
pols_tidy = 
  pols %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year","month","day")) %>% 
  mutate(month=month.name[as.numeric(month)]) 

#create a new vector president
president = c()  
for(i in 1:822){
  if(pull(pols_tidy,prez_dem)[i] == 1){
    president[i] = 'dem'
  }else{
    president[i] = 'gop'
  }
}
#binding president with pols data
pols_tidy = bind_cols(pols_tidy, president) 
colnames(pols_tidy) = c(colnames(pols_tidy[1:11]), 'president')
#final data
pols_tidy = select(pols_tidy, -day, -prez_dem, -prez_gop)

#read snp.csv
snp = read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/snp.csv")

#clean snp-month data
snp_tidy = 
 snp %>% 
  janitor::clean_names() %>% 
  mutate(year = lubridate::mdy(date)) %>%  
  separate(date, into = c("month","day","year")) %>%
 mutate(year = ifelse(year < 50, paste("20", year, sep = ""),paste("19", year, sep = ""))) %>% 
  mutate(month = month.name[as.numeric(month)]) %>% 
  relocate(year, month) %>% 
  select(-day)  

#read unemployment data
unemployment = read_csv("./data/fivethirtyeight_datasets/fivethirtyeight_datasets/unemployment.csv")

#clean unemployment data
une_tidy = 
  unemployment %>% 
   janitor::clean_names() %>% 
  pivot_longer(
    'jan':'dec',
    names_to = "month", 
    values_to = "unemploy") %>% 
  mutate(month = str_to_title(month)) %>% 
    mutate(month = match(month,month.abb)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(year = as.character(year))

#merge snp into pols
data_join = right_join(snp_tidy, pols_tidy, by = c('year','month'))

#merge unemployment into the data set
data_join2 = 
  left_join(data_join, une_tidy, by = c('year','month')) %>% 
  arrange(desc(year))


```

The `pols_tidy` data set contains 9 columns and 822 rows.It includes the number of republican and democratic governors,senators,representatives in each month through 1947 to 2015.It also captured the president's political party of each month through 1947 to 2015 using `president` variable.

The `snp_tidy` data set contains 3 columns and 787 rows. It documents the sum of the closing values of the S&P stock index of associating month from 1950 to 2015 using `close` variable.

The `une_tidy` data set contains 3 columns and 816 rows. It contains the percentage of unemployment in the associating month from 1948 to 2015 using `unemploy` variable.

The resulting data set ` data_join2` contains 11 columns and 822 rows. It is showing the closing values of the S&P stock index with `close` variable, percentage of unemployment with `unemploy` variable, number of republican and democratic governors, number of republican and democratic senators, number of republican and democratic representatives, the president's political party of each month through 1947 to 2015 using `president` variable within the `pols_tidy` data set.



# Problem 3

```{r}
#read baby names data
baby = read_csv("./data/Popular_Baby_Names.csv")

#tidy baby data set
baby_tidy = 
   baby %>% 
   janitor::clean_names() %>% 
   mutate(gender = str_to_lower(gender), ethnicity = str_to_lower(ethnicity), childs_first_name = str_to_lower(childs_first_name)) %>% 
  mutate(
    ethnicity = replace(ethnicity,ethnicity == 'asian and paci','asian and pacific islander'),
     ethnicity = replace(ethnicity,ethnicity == 'black non hisp','black non hispanic'),
    ethnicity = replace(ethnicity,ethnicity == 'white non hisp','white non hispanic') ) %>% 
  distinct()

# show Olivia's population in times
Olivia = 
  baby_tidy %>% 
  filter(childs_first_name == 'olivia') %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank") %>% 
relocate(ethnicity,'2011','2012','2013','2014','2015','2016') %>% 
 knitr::kable()

Olivia

#male popular name 
boy =
  baby_tidy %>% 
  filter(gender == 'male') %>% 
  filter(childs_first_name == "ethan") %>% 
select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
  names_from = "year_of_birth", 
  values_from = "rank") %>% 
relocate(ethnicity,'2011','2012','2013','2014','2015','2016') %>% 
 knitr::kable()
#Ethan
boy

#scatter plot 
boy_2016_df = 
  baby_tidy %>% 
  filter(year_of_birth == 2016, gender == 'male', ethnicity == 'white non hispanic') 
  
ggplot(boy_2016_df, aes(x = rank, y = count )) + 
  geom_point() +
  geom_text(label = pull(boy_2016_df,childs_first_name), nudge_x = 0.25, nudge_y = 0.25, check_overlap = T)

  
```

